### 2021-03-09 text/draft.tex
docu

The focus of this project lies primarily on practical implementation of
the prototype although care is taken to spell out design decisions in as
much detail as is needed.

@@

Even though Edoc is a public server, its database does not have a
web-ready API. In addition, since Edoc is a production server, its
underlying database cannot be used directly on pain of disturbing the
provided services. The data hence needs to be extracted from Edoc in
order to work with it. This could for example be achieved by cloning the
database of the server and running it on a local machine. However, this
route was not available due to the limited resources of the responsible
co-workers. I thus employed a workaround: Edoc has a built-in advanced
search tool where the results can be exported. Even though it is not
possible to extract all database items by default, only one of the many
available search fields has to be filled in. The complete database can
hence be extracted by only filling in the \texttt{date} field and using
a sufficiently permissive time period such as 1940 to 2020 since the
oldest record in the database was published in 1942. On January 21 2021,
this query yielded 68'345 results. These results were then exported as a
326 MB JSON file called \texttt{1942-2020.json}. \texttt{1942-2020.json}
has two drawbacks. First, the maximum file size on GitHub is 100 MB. And
second, \texttt{1942-2020.json} is too big to be handled by standard
editors requiring special editors such as Oxygen. For these reasons
\texttt{1942-2020.json} was split into smaller files that are easier
handled and can be uploaded to GitHub. For this
\texttt{Utility.split\_json} was used yielding 14 files of size 20 MB or
less containing 5000 or fewer entries each. These files are called
\texttt{raw\_master\_\{year\}.json} and saved under \texttt{/raw}.

@@

Each Edoc item has 43 main data fields. Most of these data fields belong
to the class of descriptive metadata. The data entry for an item's
descriptive metadata is undertaken by the person who uploads the item
which is in most cases one of the item's authors. The Note that an item
is first uploaded to the University of Basel's research database
(https://www.forschdb2.unibas.ch) before being automatically fed into
Edoc. There are explicit rules on how to enter the data as summarized in
the user manual (Universität Basel, n.d.). However, there is no (or at
least no systematic) manual or automatic validation of the entered data.
This means that some data fields are very heterogenous and require
further parsing (see \protect\hyperlink{gold-standard}{section ``Gold
standard''}). The data fields that are of interest to the project at
hand are discussed in more detail in (see
\protect\hyperlink{sample-data-set}{section ``Sample data set''} below).

@@

The first constraint is that each item in the sample data set needs to
have a text. Intuitively, this text could consist of a title, an
abstract, or even a full text, or any combination of the above. Similar
to manual indexing, having more information (that is, longer texts)
usually implies a higher indexing quality in machine indexing. However,
this assumption needs to be confirmed empirically for the given machine
indexing framework and data set. If indeed confirmed, it might be
advisable only to index items that have a certain minimal text length.
Therefore, in order to construct the sample data set, we require each
item to have a non-empty value in the data fields \texttt{title},
\texttt{abstract}, and \texttt{id\_number} as proxy to retrieve a full
text remotely. Note that even more context could be provided by taking
into account other data fields such as \texttt{type},
\texttt{publication} or \texttt{department}. Especially the latter might
be valuable when disambiguating homonymous or polysemous words. For
example, consider the item
\texttt{https://edoc.unibas.ch/id/eprint/76510} which is titled
``Blacking Out.'' Without further context, this title could refer
(amongst others) to a physiological phenomenon, a sudden loss of
electricity, or a measure taken in wartime. Knowing that the item was
published by a researcher employed by the Faculty of Business and
Economics (and not, say, by the Faculty of Medicine) gives us reason to
exclude the first meaning. However, this idea cannot be implemented with
the out of the box instance of Annif employed in this chapter (see
\protect\hyperlink{machine-indexing}{section ``Machine indexing''}).

The second constraint is that the index terms assigned to each item in
the sample data set must be assessable. How the assessment is conducted
in detail is discussed in \protect\hyperlink{assessment}{section
``Assessment''}. For now, it is sufficient to say that we require a
standard against which we can evaluate the assigned index terms. By
``standard'' I mean that given a vocabulary of index terms and our
sample data set, for every item in the sample data set, there is an
approved subset of the vocabulary. The production of a standard from
scratch is of course is very costly since a person, usually highly
skilled in a certain academic domain, must assign and/or approve each
item's index terms (!!source). For this reason we try to avoid having to
produce a standard. One way of doing so is by requiring items in the
sample data set to have non-empty \texttt{keywords} data field. The
caveats of this approach are discussed in
\protect\hyperlink{assessment}{section ``Assessment''}.

Taking into account the third constraint simply means that we do not
employ any further restrictions. We can hence construct the sample data
set by choosing exactly those items from \texttt{/raw} which have
non-empty data fields \texttt{title}, \texttt{abstract},
\texttt{id\_number}, and \texttt{keywords}. We do so by calling
\texttt{Data.select\_from\_file} iteratively for all files in
\texttt{/raw}. The resulting file is saved in \texttt{/selected} as
\texttt{selected\_master.json}.

@@

Of the 68'345 items in \texttt{/raw}, all have a title (non-empty
\texttt{title} data field), a little more than half of the items have an
abstract (37'381 items with non-empty \texttt{abstract} data field),
roughly half of the items have an ID (35'355 items with non-empty
\texttt{id\_number} data field),\footnote{Note that when using the facet
  by blank on \texttt{id\_number} in OpenRefine, there are 57'153
  matches; this is due to the fact that many items with an ID such as
  DOI have secondary or tertiary IDs such as ISI or PMID.} and less than
10\% of the items have keywords (6'660 items with non-empty
\texttt{keywords} data field). The sample data set as requires all the
above data fields to be non-empty; \texttt{/sample/sample\_master.json}
has 4'111 items and hence constitutes 6\% of the raw data.

@@

Annif (Suominen et al. 2021) is an open source multi-algorithm automated
indexing tool developed at the Finnish National Library (see Suominen
2019 for an overview). Training a local Annif instance for machine
indexing is currently evaluated at different libraries, for example at
the German National Library (Nagelschmidt 12.11.2020). As stated in
section \protect\hyperlink{introduction}{section ``Introduction''}, in
this project I am interested in using Annif out of the box. This means
that instead of training a local Annif instance, I will be relying on
the Annif REST-style API (see https://api.annif.org/).\footnote{This API
  is specifically intended for testing. For integration into a
  production system, the Finto AI API is available (see
  https://ai.finto.fi/v1/ui/).} The advantage of this approach is that
it is quick and relatively easy to implement. The disadvantage is that
no custom algorithms or vocabularies can be used. The task of
constructing a gold standard (see
\protect\hyperlink{gold-standard}{section ``Gold standard''}) remains
the same.

The vocabularies provided by the Annif API are the General Finnish
Ontology or YSO in short (see https://finto.fi/yso/en/) and Wikidata
(see https://www.wikidata.org/). Each vocabulary is combined with an
algorithm or algorithm ensemble respectively and a specific language to
create a so-called Annif ``project.'' The five projects used in what
follows are summarized in Table 1. A high-level description of the
algorithm(s) is given by Suominen (2019, 7--10). Information on
implementation can be found at the Annif GitHub Wiki (see
https://github.com/NatLibFi/Annif/wiki).

@@

Machine indexing with Annif was implemented using the annif-client
Python library (see https://pypi.org/project/annif-client/). The desired
output is that every item in the Edoc sample data set is assigned
subject terms by all available Annif API projects (namely,
\texttt{yso-en}, \texttt{yso-maui-en}, \texttt{yso-bonsai-en},
\texttt{yso-fasttext-en}, \texttt{wikidata-en}) based on all available
text bases (title or title and abstract). In order to distinguish these
assignments, we use a unique ID (called ``marker'' in what follows) for
each Annif configuration. A marker is constructed by the convention:
\texttt{\{project\_id\}-\{abstract\}-\{fulltext\}-\{n\}-\{threshold\}},
where \texttt{project\_id} is the Annif project, \texttt{abstract} and
\texttt{fulltext} are boolean variables, \texttt{n} is the maximum
number of suggestions per item, and \texttt{threshold} is a variable for
a threshold Annif score.

In order to generate the ouput, we call
\texttt{Data.super\_enrich\_with\_annif}, thereby iteratively calling
\texttt{Data.enrich\_with\_annif} for all items in
\texttt{/sample/sample\_master.json} for all Annif projects. The output
is saved in \texttt{/indexed/} as \texttt{indexed\_master.json}.

@@

As explained in \protect\hyperlink{machine-indexing}{section ``Machine
indexing''}, Annif assigns index terms from a controlled vocabulary. If
we want to assess the quality of the indexing via a gold standard, we
must therefore ensure that the gold standard makes use of the vocabulary
used by Annif. The relevant (English) vocabularies are Wikidata and YSO.
The next step in constructing the derivative gold standard is hence to
match the extracted and cleaned keywords with keywords from Wikidata and
YSO. This process is called ``reconciliation'' (see
https://docs.openrefine.org/manual/reconciling) and the tool of choice
for this task is OpenRefine).\footnote{For data refinement I use
  OpenRefine version 3.4.1 (https://openrefine.org/). Data manipulation
  in OpenRefine is tracked: the manipulation history of some data can be
  exported as JSON file and then reproduced (on the same data on a
  different machine or on different data) by loading said file. I will
  supply a corresponding manipulation history whenever appropriate. Note
  that when using OpenRefine with larger files (and there are many such
  files in this project), the memory allocation to OpenRefine must be
  manually increased (see
  https://docs.openrefine.org/manual/installing/\#increasing-memory-allocation
  for details). Also note that OpenRefine always counts the number of
  records by the first column (a fact which caused me many headaches).}

@@

Second, systematic biases were identified and removed. The most
prevalent bias was an due to a suggestion's type. As stated above, the
reconciliation service API was not constrained by type but had access to
the complete Wikidata database. Since Wikidata is an ontology that
encompasses everything, it also features types whose concepts cannot
qualify as subject terms (at least in the present context). The most
prominent example is the type Q13442814 ``scholarly article.'' Wikidata
contains the metadata of many scholarly articles. Now, for some of our
keywords, there is a scholarly article with a title that exactly matches
the keyword; and since there is no restriction concerning the type, the
scholarly article is suggested with high confidence (see
https://github.com/OpenRefine/OpenRefine/wiki/Reconciliation-Service-API
for details). To generalize, suggestions with types whose concepts are
proper names are usually incorrect. Based on this observation,
suggestions with types such as ``scholarly article,'' ``clinical
trial,'' and ``scientific journal'' were rejected.\footnote{The complete
  list includes ``academic journal,'' ``open access journal,''
  ``thesis,'' ``doctoral thesis,'' and ``natural number.''} Suggestions
types such as ``human,'' ``album,'' and ``commune of France'' were
manually verified (i.e., checked for correctness).\footnote{The complete
  list includes ``film,'' ``musical group,'' ``business,'' ``literary
  work,'' ``television series,'' ``organization,'' ``family name,''
  ``written work,'' ``video game,'' ``single, television series
  episode,'' ``painting,'' , ``city of the United States,''
  ``magazine,'' ``studio album,'' ``year,'' ``nonprofit organization,''
  ``border town,'' ``international organization,'' ``political party,''
  ``software,'' ``song,'' ``website,'' ``article comic strip,''
  ``collection,'' ``commune of Italy,'' ``fictional human,'' ``film,''
  ``government agency,'' ``village,'' ``academic journal article,''
  ``female given name,'' and ``poem.''}

With these improvements in place, each keyword with a suggestion was
assigned a QID via the ``Add cloumns from reconciled values''-function
(and similar for YSO and MeSH identifiers). The data was then exported
and saved as \texttt{/keywords/keywords\_reference.json}. Keywords with
a QID now constitute 69\% of all keywords and 78\% of their total
occurrences in the sample data set, but these numbers are significantly
lower for the MeSH identifier (53.6\% of all keywords with only 28.8\%
of all occurrences) and especially low for the YSO identifier (26.9\% of
all keywords with 11\% of all occurrences). In both cases, the problem
is due to the fact that Wikidata's mapping of MeSH respectively YSO is
only partial. There can be two reasons for this state of affairs for a
given entry: either there is no match between Wikidata and YSO or MeSH
(after all, Wikidata is much larger than MeSH and YSO taken together),
or there is a match but it has not yet been added to the mapping. In the
latter case, at least with respect to YSO, there is a solution: Finto
provides a REST-style API to access the YSO vocabulary directly (see
https://api.finto.fi/). For each keyword in the list of reference
keywords that lacks a YSO identifier, the Finto API is queried; if a
term turns up, it is added to the keyword. The effect of this second
reconcilement is detailed in Figure 4. It is achieved by calling
\texttt{Keywords.enrich\_with\_yso} on
\texttt{/keywords/keywords\_reference.json} and saving the output as
\texttt{/keywords/keywords\_reference\_master.json}

@@

The metrics used for the assessment are precision, recall and F1-score.
Precision and recall are standard metrics for indexing quality (e.g.,
Gantert 2016, 197) whereby the F1 score plays are more prominent role in
the assessment of machine indexing (e.g., Suominen 2019, 11--14; Toepfer
and Kempf 2016, 93f.). Of course, there is a host of alternative metrics
(such as indexing consistency, transparency, reproducability) that are
neglected here.

@@

A surprising result is that there was no performance gain achieved by
increasing the text basis from titles to titles and abstracts (see
\protect\hyperlink{creating-the-data-foundation}{section ``Creating the
data foundation''}): if anything, the performance for the extended text
basis was slightly worse (see Figure 6, right). There is again no easy
explanation for this find since there are multiple plausible causes
which cannot be distangled here. For example, the abstract of an item in
the Edoc sample data set might be less representative of the item's
content than its title. In that case, the observed worse perfomance of
the Annif configurations based on title and abstracts would even be
expected. Alternatively, it might be the case that the problem lies with
Annif: a larger text basis introduces more confounding factors and might
make it harder to extract the subject terms required. An experimentum
crucis to decide between the two discussed explanations would be to
further extend the text basis to include full texts in addition to
titles and abstracts.

@@

As mentioned in \protect\hyperlink{introduction}{section
``Introduction''}, all files, data, and code used in this project is
available at https://github.com/MHindermann/mas and will not be
replicated here.

>>>>

The focus of this project lies primarily on practical implementation of
the prototype although care is taken to spell out design decisions in as
much detail as is needed. All code written and data generated during the
process of implementing the prototype is available at
\url{https://github.com/MHindermann/mas}:

\begin{itemize}
\tightlist
\item
  Text and images available at
  \href{https://github.com/MHindermann/mas/tree/main/text}{\texttt{/text/}}
\item
  Codebase and data available at
  \href{https://github.com/MHindermann/mas/tree/main/files}{\texttt{/files/}}
\item
  Documentation available at
  \href{https://github.com/MHindermann/mas/tree/main/docs}{\texttt{/docs/}}
\end{itemize}

@@

Even though Edoc is a public server, its database does not have a
web-ready API. In addition, since Edoc is a production server, its
underlying database cannot be used directly on pain of disturbing the
provided services. The data hence needs to be extracted from Edoc in
order to work with it. This could for example be achieved by cloning the
database of the server and running it on a local machine. However, this
route was not available due to the limited resources of the responsible
co-workers. I thus employed a workaround: Edoc has a built-in advanced
search tool where the results can be exported. Even though it is not
possible to extract all database items by default, only one of the many
available search fields has to be filled in. The complete database can
hence be extracted by only filling in the \texttt{date} field and using
a sufficiently permissive time period such as 1940 to 2020 since the
oldest record in the database was published in 1942. On January 21 2021,
this query yielded 68'345 results. These results were then exported as a
326 MB JSON file called \texttt{1942-2020.json}. \texttt{1942-2020.json}
has two drawbacks. First, the maximum file size on GitHub is 100 MB. And
second, \texttt{1942-2020.json} is too big to be handled by standard
editors requiring special editors such as Oxygen. For these reasons
\texttt{1942-2020.json} was split into smaller files that are easier
handled and can be uploaded to GitHub. For this
\texttt{Utility.split\_json} was used yielding 14 files of size 20 MB or
less containing 5000 or fewer entries each. These files are called
\texttt{raw\_master\_\{year\}.json} and saved under
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}}.

@@

Each Edoc item has 43 main data fields. Most of these data fields belong
to the class of descriptive metadata. The data entry for an item's
descriptive metadata is undertaken by the person who uploads the item
which is in most cases one of the item's authors. The Note that an item
is first uploaded to the University of Basel's research database
\url{https://www.forschdb2.unibas.ch} before being automatically fed
into Edoc. There are explicit rules on how to enter the data as
summarized in the user manual (Universität Basel, n.d.). However, there
is no (or at least no systematic) manual or automatic validation of the
entered data. This means that some data fields are very heterogenous and
require further parsing (see \protect\hyperlink{gold-standard}{section
``Gold standard''}). The data fields that are of interest to the project
at hand are discussed in more detail in (see
\protect\hyperlink{sample-data-set}{section ``Sample data set''} below).

@@

The first constraint is that each item in the sample data set needs to
have a text. Intuitively, this text could consist of a title, an
abstract, or even a full text, or any combination of the above. Similar
to manual indexing, having more information (that is, longer texts)
usually implies a higher indexing quality in machine indexing. However,
this assumption needs to be confirmed empirically for the given machine
indexing framework and data set. If indeed confirmed, it might be
advisable only to index items that have a certain minimal text length.
Therefore, in order to construct the sample data set, we require each
item to have a non-empty value in the data fields \texttt{title},
\texttt{abstract}, and \texttt{id\_number} as proxy to retrieve a full
text remotely. Note that even more context could be provided by taking
into account other data fields such as \texttt{type},
\texttt{publication} or \texttt{department}. Especially the latter might
be valuable when disambiguating homonymous or polysemous words. For
example, consider the item
\href{https://edoc.unibas.ch/id/eprint/76510}{\texttt{https://edoc.unibas.ch/id/eprint/76510}}
which is titled ``Blacking Out.'' Without further context, this title
could refer (amongst others) to a physiological phenomenon, a sudden
loss of electricity, or a measure taken in wartime. Knowing that the
item was published by a researcher employed by the Faculty of Business
and Economics (and not, say, by the Faculty of Medicine) gives us reason
to exclude the first meaning. However, this idea cannot be implemented
with the out of the box instance of Annif employed in this chapter (see
\protect\hyperlink{machine-indexing}{section ``Machine indexing''}).

The second constraint is that the index terms assigned to each item in
the sample data set must be assessable. How the assessment is conducted
in detail is discussed in \protect\hyperlink{assessment}{section
``Assessment''}. For now, it is sufficient to say that we require a
standard against which we can evaluate the assigned index terms. By
``standard'' I mean that given a vocabulary of index terms and our
sample data set, for every item in the sample data set, there is an
approved subset of the vocabulary. The production of a standard from
scratch is of course is very costly since a person, usually highly
skilled in a certain academic domain, must assign and/or approve each
item's index terms. For this reason we try to avoid having to produce a
standard. One way of doing so is by requiring items in the sample data
set to have non-empty \texttt{keywords} data field. The caveats of this
approach are discussed in \protect\hyperlink{assessment}{section
``Assessment''}.

Taking into account the third constraint simply means that we do not
employ any further restrictions. We can hence construct the sample data
set by choosing exactly those items from
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}}
which have non-empty data fields \texttt{title}, \texttt{abstract},
\texttt{id\_number}, and \texttt{keywords}. We do so by calling
\texttt{Data.select\_from\_file} iteratively for all files in
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}}.
The resulting file is saved in
\href{https://github.com/MHindermann/mas/tree/main/files/sample}{\texttt{/sample/}}
as \texttt{sample\_master.json}.

@@

Of the 68'345 items in
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}},
all have a title (non-empty \texttt{title} data field), a little more
than half of the items have an abstract (37'381 items with non-empty
\texttt{abstract} data field), roughly half of the items have an ID
(35'355 items with non-empty \texttt{id\_number} data field),\footnote{Note
  that when using the facet by blank on \texttt{id\_number} in
  OpenRefine, there are 57'153 matches; this is due to the fact that
  many items with an ID such as DOI have secondary or tertiary IDs such
  as ISI or PMID.} and less than 10\% of the items have keywords (6'660
items with non-empty \texttt{keywords} data field). The sample data set
as requires all the above data fields to be non-empty;
\texttt{/sample/sample\_master.json} has 4'111 items and hence
constitutes 6\% of the raw data.

@@

Annif (Suominen et al. 2021) is an open source multi-algorithm automated
indexing tool developed at the Finnish National Library (see Suominen
2019 for an overview). Training a local Annif instance for machine
indexing is currently evaluated at different libraries, for example at
the German National Library (Nagelschmidt 12.11.2020). As stated in
section \protect\hyperlink{introduction}{section ``Introduction''}, in
this project I am interested in using Annif out of the box. This means
that instead of training a local Annif instance, I will be relying on
the Annif REST-style API (see \url{https://api.annif.org/}).\footnote{This
  API is specifically intended for testing. For integration into a
  production system, the Finto AI API is available (see
  \url{https://ai.finto.fi/v1/ui/}).} The advantage of this approach is
that it is quick and relatively easy to implement. The disadvantage is
that no custom algorithms or vocabularies can be used. The task of
constructing a gold standard (see
\protect\hyperlink{gold-standard}{section ``Gold standard''}) remains
the same.

The vocabularies provided by the Annif API are the General Finnish
Ontology or YSO in short (see \url{https://finto.fi/yso/en/}) and
Wikidata (see \url{https://www.wikidata.org/}). Each vocabulary is
combined with an algorithm or algorithm ensemble respectively and a
specific language to create a so-called Annif ``project.'' The five
projects used in what follows are summarized in Table 1. A high-level
description of the algorithm(s) is given by Suominen (2019, 7--10).
Information on implementation can be found at the Annif GitHub Wiki (see
\url{https://github.com/NatLibFi/Annif/wiki}).

@@

Machine indexing with Annif was implemented using the annif-client
Python library (see \url{https://pypi.org/project/annif-client/}). The
desired output is that every item in the Edoc sample data set is
assigned subject terms by all available Annif API projects (namely,
\texttt{yso-en}, \texttt{yso-maui-en}, \texttt{yso-bonsai-en},
\texttt{yso-fasttext-en}, \texttt{wikidata-en}) based on all available
text bases (title or title and abstract). In order to distinguish these
assignments, we use a unique ID (called ``marker'' in what follows) for
each Annif configuration. A marker is constructed by the convention:
\texttt{\{project\_id\}-\{abstract\}-\{fulltext\}-\{n\}-\{threshold\}},
where \texttt{project\_id} is the Annif project, \texttt{abstract} and
\texttt{fulltext} are boolean variables, \texttt{n} is the maximum
number of suggestions per item, and \texttt{threshold} is a variable for
a threshold Annif score.

In order to generate the ouput, we call
\texttt{Data.super\_enrich\_with\_annif}, thereby iteratively calling
\texttt{Data.enrich\_with\_annif} for all items in
\texttt{/sample/sample\_master.json} for all Annif projects. The output
is saved in
\href{https://github.com/MHindermann/mas/tree/main/files/indexed}{\texttt{/indexed/}}
as \texttt{indexed\_master.json}.

@@

As explained in \protect\hyperlink{machine-indexing}{section ``Machine
indexing''}, Annif assigns index terms from a controlled vocabulary. If
we want to assess the quality of the indexing via a gold standard, we
must therefore ensure that the gold standard makes use of the vocabulary
used by Annif. The relevant (English) vocabularies are Wikidata and YSO.
The next step in constructing the derivative gold standard is hence to
match the extracted and cleaned keywords with keywords from Wikidata and
YSO. This process is called ``reconciliation'' (see
\url{https://docs.openrefine.org/manual/reconciling}) and the tool of
choice for this task is OpenRefine).\footnote{For data refinement I use
  OpenRefine version 3.4.1 (see \url{https://openrefine.org/}). Data
  manipulation in OpenRefine is tracked: the manipulation history of
  some data can be exported as JSON file and then reproduced (on the
  same data on a different machine or on different data) by loading said
  file. I will supply a corresponding manipulation history whenever
  appropriate. Note that when using OpenRefine with larger files (and
  there are many such files in this project), the memory allocation to
  OpenRefine must be manually increased (see
  \url{https://docs.openrefine.org/manual/installing/\#increasing-memory-allocation}
  for details). Also note that OpenRefine always counts the number of
  records by the first column (a fact which caused me many headaches).}

@@

Second, systematic biases were identified and removed. The most
prevalent bias was an due to a suggestion's type. As stated above, the
reconciliation service API was not constrained by type but had access to
the complete Wikidata database. Since Wikidata is an ontology that
encompasses everything, it also features types whose concepts cannot
qualify as subject terms (at least in the present context). The most
prominent example is the type Q13442814 ``scholarly article.'' Wikidata
contains the metadata of many scholarly articles. Now, for some of our
keywords, there is a scholarly article with a title that exactly matches
the keyword; and since there is no restriction concerning the type, the
scholarly article is suggested with high confidence (see
\url{https://github.com/OpenRefine/OpenRefine/wiki/Reconciliation-Service-API}
for details). To generalize, suggestions with types whose concepts are
proper names are usually incorrect. Based on this observation,
suggestions with types such as ``scholarly article,'' ``clinical
trial,'' and ``scientific journal'' were rejected.\footnote{The complete
  list includes ``academic journal,'' ``open access journal,''
  ``thesis,'' ``doctoral thesis,'' and ``natural number.''} Suggestions
types such as ``human,'' ``album,'' and ``commune of France'' were
manually verified (i.e., checked for correctness).\footnote{The complete
  list includes ``film,'' ``musical group,'' ``business,'' ``literary
  work,'' ``television series,'' ``organization,'' ``family name,''
  ``written work,'' ``video game,'' ``single, television series
  episode,'' ``painting,'' , ``city of the United States,''
  ``magazine,'' ``studio album,'' ``year,'' ``nonprofit organization,''
  ``border town,'' ``international organization,'' ``political party,''
  ``software,'' ``song,'' ``website,'' ``article comic strip,''
  ``collection,'' ``commune of Italy,'' ``fictional human,'' ``film,''
  ``government agency,'' ``village,'' ``academic journal article,''
  ``female given name,'' and ``poem.''}

With these improvements in place, each keyword with a suggestion was
assigned a QID via the ``Add cloumns from reconciled values''-function
(and similar for YSO and MeSH identifiers). The data was then exported
and saved as \texttt{/keywords/keywords\_reference.json}. Keywords with
a QID now constitute 69\% of all keywords and 78\% of their total
occurrences in the sample data set, but these numbers are significantly
lower for the MeSH identifier (53.6\% of all keywords with only 28.8\%
of all occurrences) and especially low for the YSO identifier (26.9\% of
all keywords with 11\% of all occurrences). In both cases, the problem
is due to the fact that Wikidata's mapping of MeSH respectively YSO is
only partial. There can be two reasons for this state of affairs for a
given entry: either there is no match between Wikidata and YSO or MeSH
(after all, Wikidata is much larger than MeSH and YSO taken together),
or there is a match but it has not yet been added to the mapping. In the
latter case, at least with respect to YSO, there is a solution: Finto
provides a REST-style API to access the YSO vocabulary directly (see
\url{https://api.finto.fi/}). For each keyword in the list of reference
keywords that lacks a YSO identifier, the Finto API is queried; if a
term turns up, it is added to the keyword. The effect of this second
reconcilement is detailed in Figure 4. It is achieved by calling
\texttt{Keywords.enrich\_with\_yso} on
\texttt{/keywords/keywords\_reference.json} and saving the output as
\texttt{/keywords/keywords\_reference\_master.json}

@@

The metrics used for the assessment are precision, recall and F1-score.
Precision and recall are standard metrics for indexing quality (e.g.,
Gantert 2016, 197) whereby the F1 score plays are more prominent role in
the assessment of machine indexing (e.g., Suominen 2019, 11--14; Toepfer
and Kempf 2016, 93f.). Of course, there is a host of alternative metrics
(such as indexing consistency, transparency, reproducibility) that are
neglected here.

@@

A surprising result is that there was no performance gain achieved by
increasing the text basis from titles to titles and abstracts (see
\protect\hyperlink{creating-the-data-foundation}{section ``Creating the
data foundation''}): if anything, the performance for the extended text
basis was slightly worse (see Figure 6, right). There is again no easy
explanation for this find since there are multiple plausible causes
which cannot be distangled here. For example, the abstract of an item in
the Edoc sample data set might be less representative of the item's
content than its title. In that case, the observed worse performance of
the Annif configurations based on title and abstracts would even be
expected. Alternatively, it might be the case that the problem lies with
Annif: a larger text basis introduces more confounding factors and might
make it harder to extract the subject terms required. An experimentum
crucis to decide between the two discussed explanations would be to
further extend the text basis to include full texts in addition to
titles and abstracts.

@@

As mentioned in \protect\hyperlink{introduction}{section
``Introduction''}, all files, data, and code used in this project is
available at \url{https://github.com/MHindermann/mas} and will not be
replicated here.
